Model Selection
================

This is a R Markdown document. Markdown is a simple formatting syntax for authoring web pages,
and a very nice way of distributing an analysis. It has some very simple syntax rules.


```{r}
library(ISLR)
summary(Hitters)
```
There are some missing values here, so before we proceed we will remove them:

```{r}
Hitters=na.omit(Hitters)
with(Hitters,sum(is.na(Salary)))




Best Subset regression
------------------------
WE will now use the package 'leaps' to evaluate all the best-subset models. #best subset model looks for best model (lowest RSS) at each size
```{r}
library(leaps)
regfit.full=regsubsets(Salary~.,data=Hitters) #finds best subset of each size
summary(regfit.full) 
```
It gives by default best-subsets up to size 8;lets increase that to 19,
i.e. all the variables
```{r}
regfit.full=regsubsets(Salary~.,data=Hitters,nvmax=19)
reg.summary=summary(regfit.full)
names(reg.summary)
plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp")
which.min(reg.summary$cp)
points(10,reg.summary$cp[10],pch=20,col="red") #colors 10th point red on plot

There is a plot method for the 'regsubsets' object
```{r}
plot(regfit.full,scale="Cp")
coef(regfit.full,10)
```




Forward Stepwise Selection
--------------------------
Here we use the 'regsubsets' function but specify the 'method="forward" option:
```{r}
regfit.fwd=regsubsets(Salary~.,data=Hitters,nvmax=19,method="forward")
summary(regfit.fwd)
plot(regfit.fwd,scale="Cp")
```




Model Selection Using a Validation Set
---------------------------------------
Let's make a training and validation set, so that we can chose a good subset model.
We will do it using a slightly different approach from what was done 
in the book.
```{r}
dim(Hitters)
set.seed(1)
train=sample(seq(263),180,replace=FALSE) #use approximately 2/3 of 263 rows
train
regfit.fwd=regsubsets(Salary~.,data=Hitters[train,],nvmax=19,method="forward")
```
Now we will make predictions on the observations not used for training.
We know now there are 19 models, so we set up some vectors to record the errors. We have to do a bit of work here, because there is no predict method for 'regsubsets'.
```{r}
val.errors=rep(NA,19)
x.test=model.matrix(Salary~.,data=Hitters[-train,]
for(i in 1:19){
  coefi=coef(regfit.fwd,id=i)
  pred=x.test[,names(coefi)]%*%coefi
  val.errors[i]=mean((Hitters$Salary[-train]-pred)^2)
}
plot(sqrt(val.errors),ylab="Root MSE",ylim=c(300,400),pch=19,type="b")
points(sqrt(regfit.fwd$rss[-1]/180),col="blue",pch=19,type="b")
legend("topright",legend=c("Training","Validation"),col=c("blue","black"),pch=19)
```
As we expect, the training error goes down monotonically as the model gets bigger, but not so 
 for the validation error.

This was a little tedious - not having a predict method for 'regsubsets'. So we will write one!
```{r}
predict.regsubsets=function(object,newdata,id,...){
  form=as.formula(object$call[[2]])
  mat=model.matrix(form,newdata)
  coefi=coef(object,id=id)
  mat[,names(coefi)]%*%coefi
}
```
