Nonlinear Models
==========================
Here we explore the use of nonlinear models using some tools in R

```{r}
require(ISLR)
attach(Wage)
```

Polynomials
------------

First we will use polynomials, and focus on a single predictor age:

```{r}
fit=lm(wage~poly(age,4),data=Wage)
summary(fit)
```

The `poly()` function generates a basis of *othogonal polynomials* (variables are uncorrelated)
Let's make a plot of the fitted function, along with the standard errors of the fit.

```{r fig.width=7,fig.height=6}
agelims=range(age)
age.grid=seq(from=agelims[1],to=agelims[2])
preds=predict(fit,newdata=list(age=age.grid),se=TRUE)
se.bands=cbind(preds$fit+2*preds$se,preds$fit-2*preds$se)
plot(age,wage,col="darkgrey")
lines(age.grid,preds$fit,lwd=2,col="blue")
matlines(age.grid,se.bands,col="blue",lty=2)
```

There are other more direct ways of doing this in R. For example

```{r}
fita=lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
summary(fita)
```

Here `I()` is a *wrapper* function; we need it because `age^2` means something to the formula language,
while `I(age^2)` is protected.
The coefficients are different to those we got before! However, the fits are the same:

```{r}
plot(fitted(fit),fitted(fita))
```

By using orthogonal polynomials in this simple way, it turns out that we can separately test
for each coefficient. So if we look at the summary again, we can see that the linear, quadratic
and cubic terms are significant, but not the quartic.

```{r}
summary(fit)
```

This only works with linear regression, and if there is a single predictor. IN general we would use `anova()`
as this next example demonstrates

```{r}
fita=lm(wage~education,data=Wage)
fitb=lm(wage~education+age,data=Wage)
fitc=lm(wage~education+poly(age,2),data=Wage)
fitd=lm(wage~education+poly(age,3),data=Wage)
anova(fita,fitb,fitc,fitd) #gives output and compares models to see which variabels are significnat (by examining f-statistic)

```

###Polynomial logistic regression

Now we fit a logistic regression model to a binary response variable,
constructed from `wage`. We code the big earners (`>250k`) as 1, else 0.

```{r}
fit=glm(I(wage>250) ~ poly(age,3), data=Wage, family=binomial)
summary(fit)
preds=predict(fit,list(age=age.grid),se=T)
se.bands=preds$fit+cbind(fit=0,lower=-2*preds$se,upper=2*preds$se)
se.bands[1:5,]
```

We have done the computation on the logit scale. To transform we need to apply the inverse logit
mapping
$$p=\frac{e^\eta}{1+e^\eta}.$$
(Here we have used the ability of MarkDown to interpret TeX expressions.)
We can do this simultaneously for all three columns of `se.bands`:

```{r}
prob.bands=exp(se.bands)/(1+exp(se.bands))
matplot(age.grid,prob.bands,col="blue",lwd=c(2,1,1),lty=c(1,2,2),type="l",ylim=c(0,.1))
points(jitter(age),I(wage>250)/10,pch="I",cex=.5)
```
